{\fontfamily{ptm}\selectfont A Parameter Free Accelerated Proximal Gradient Method Without Restarting }

### 2025-01-23 Notebook LaTEX/R-WAPG/paper3.tex
>w<, big fixes and changes

@@

        For $x \in \dom g$, by definition we have: 

@@

        The model function $\widetilde{\mathcal M}^{L^{-1}}(\cdot; y)$ is an $L + \mu$ strongly convex function, so it admits quadratic growth conditions over minimizer $Ty$ and for all $x \in \RR^n$: 

@@

@@

            + D_f(Ty, y). 

@@

        Now, consider the lower bound of $D_f(x, y) \ge \mu/2\Vert x - y\Vert^2$ which is true for all $x, y \in \RR^n$ by strong convexity of $f$. 
        Then it has for all $x \in \RR^n, y \in \RR^n$: 

@@

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\
        Assume $0 \le \mu < L$.
        Fix any $k \in \mathbb Z_+$. 
        For any $(v_k, x_k), \alpha_k \in (0, 1), \gamma_k > 0$, let $\hat \gamma_{k + 1}$, and vectors $y_k, v_{k + 1}, x_{k + 1}$ be given by: 
        \begin{align}
            \hat \gamma_{k + 1} &= (1 - \alpha_k)\gamma_k + \mu \alpha_k, \label{eqn:stepwise-wapg-eqn1}
            \\
            y_k &= 
            (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k), \label{eqn:stepwise-wapg-eqn2}
            \\
            g_k &= \mathcal G_L y_k, \label{eqn:stepwise-wapg-eqn3}
            \\
            v_{k + 1} &= \hat\gamma^{-1}_{k + 1}
            (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), \label{eqn:stepwise-wapg-eqn4}
            \\
            x_{k + 1} &= T_L y_k. \label{eqn:stepwise-wapg-eqn5}
        \end{align}
    \end{definition}
    \begin{observation}\label{obs:stepwise-wapg}
        Let's observe Definition \ref{def:stepwise-wapg} closely. 
        We make two crucial observations here. 
        We have $x_k, y_k, v_k$ lie on the same line because and they have the following equivalent equalities: 
        \begin{align*}
            y_k - v_k &= 
            \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k}(x_k - y_k), 
            \tag{Q1}\label{eqn:Q1}
            \\
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
            \tag{Q2}\label{eqn:Q2}
        \end{align*}
    \end{observation}
        To see \eqref{eqn:Q1}, observe 
        \begin{align*}
            y_k - v_k &= 
            \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k}(x_k - y_k)
            \\
            \iff 
            -(\alpha_k \gamma_k \hat \gamma^{-1}_{k + 1} + 1)y_k
            &= 
            - \alpha_k \gamma_k \hat \gamma^{-1}_{k + 1}v_k - x_k
            \\
            \iff 
            y_k &= 
            \frac{
                \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}v_k + x_k
            }{1 + \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}}
            \\
            &=  
            \frac{\alpha_k \gamma_k v_k + \hat \gamma_{k + 1} x_k}{\gamma_k + \alpha_k \mu}.
        \end{align*}
        On the first equality \eqref{eqn:Q1}, we multiplied both side of the equation by $\hat\gamma_{k + 1}/\alpha_k \gamma_k$ and then grouped $y_k$. 
        The last equality comes by multiplying both the numerator and denominator by $\hat \gamma_{k + 1}$, then simplifies the denominator by rearranging \eqref{eqn:stepwise-wapg-eqn1} into $\hat \gamma_{k + 1} + \alpha_k \gamma_k = \gamma_k + \alpha_k \mu$. 
        To see the second equality \eqref{eqn:Q2}, consider \eqref{eqn:stepwise-wapg-eqn2}: 
        \begin{align*}
            y_k &= (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k)
            \\
            \iff
            y_k - x_k &= 
            (\gamma_k + \alpha_k \mu)^{-1}
            (\alpha_k \gamma_k v_k - (\gamma_k + \alpha_k \mu)x_k + \hat \gamma_{k + 1} x_k)
            \\
            \iff 
            (\gamma_k + \alpha_k \mu)(y_k - x_k)
            &= 
            \alpha_k\gamma _kv_k + 
            (\hat \gamma_k - \gamma_k - \alpha_k \mu) x_k
            \\
            &= \alpha_k \gamma_k v_k - \alpha_k \gamma_k x_k 
            \\
            &= \alpha_k \gamma_k(v_k - x_k)
            \\
            \iff 
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
        \end{align*}
        On the second equality that follows, we substituted $\hat\gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \alpha_k \mu$ given by \eqref{eqn:stepwise-wapg-eqn1}. 
        Therefore, $x_k, v_k, y_k$ lies on the same line because \eqref{eqn:Q1} indicates $y_k - v_k$ parallels to $-(y_k - x_k)$ and both vector shares the same head which anchors at $y_k$.

@@

        The proof is direct algebra: 

@@

        We started with $F(x_{k + 1}) + R_{k + 1} \iff \eqref{eqn:stepwise-lya-proof-eqn1}$, 
        then $\hat \gamma/2\Vert v_{k + 1} - x^*\Vert^2 = \eqref{expr:stepwise-lya-expr2.2}$ via \eqref{eqn:stepwise-lya-proof-eqn2}. 
        We add (\ref{eqn:stepwise-lya-proof-eqn1}), (\ref{expr:stepwise-lya-expr2.2}) to get \eqref{eqn:stepwise-lya-proof-eqn3.1}. 
        Finally, the entire LHS of \eqref{ineq:stepwise-lya-2} equals to the RHS of \eqref{eqn:stepwise-lya-proof-eqn3.2}. 
        Going from \eqref{eqn:stepwise-lya-proof-eqn3.2} to \eqref{ineq:stepwise-lya-proof-ineq4}, we used Theorem \ref{thm:prox-grad-ineq} which yields the desired inequality. 

@@

        \label{eqn:stepwise-lya-proof-eqn2}

@@

>>>>

        \textbf{The remainder of this paper is organized as follows. }

@@

        The proof is direct using algebra and the definitions. 
        For $x \in \dom g, y \in \RR^n$, using the definition of $\widetilde{\mathcal M}^{L^{-1}}(x; y)$ it has: 

@@

        The proof is direct algebra and it starts by deriving an intermediate inequality from the quadratic growth condition of the model function.
        Then it uses the characterization of strong convexity to simplify it to obtain the desired results. 
        $\widetilde{\mathcal M}^{L^{-1}}(\cdot; y)$ is $L + \mu$ strongly convex by Assumption \ref{ass:smooth-nsmooth-additive} and its definition. 
        We see that it admits quadratic growth over the minimizer $Ty$ therefore for all $x \in \RR^n$ it follows that: 

@@

            \quad 
            \text{By Lemma \ref{lemma:pg-envelope}}

@@

            + D_f(Ty, y). \quad \text{By $\mu \ge 0$}

@@

        From $\mu \ge 0$ strong convexity of $f$, it has $D_f(x, y) \ge \mu/2\Vert x - y\Vert^2$ which is true for all $x, y \in \RR^n$. 
        Consequently, it has for all $x \in \RR^n, y \in \RR^n$: 

@@

    \begin{definition}[Stepwise weak accelerated proximal gradient]\label{def:stepwise-wapg}\;\\
        Assume $0 \le \mu < L$.
        Fix any $k \in \mathbb Z_+$. 
        For any $(v_k, x_k), \alpha_k \in (0, 1), \gamma_k > 0$, let $\hat \gamma_{k + 1}$, and vectors $y_k, v_{k + 1}, x_{k + 1}$ be given by: 
        \begin{align}
            \hat \gamma_{k + 1} &= (1 - \alpha_k)\gamma_k + \mu \alpha_k, \label{eqn:stepwise-wapg-eqn1}
            \\
            y_k &= 
            (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k), \label{eqn:stepwise-wapg-eqn2}
            \\
            g_k &= \mathcal G_L y_k, \label{eqn:stepwise-wapg-eqn3}
            \\
            v_{k + 1} &= \hat\gamma^{-1}_{k + 1}
            (\gamma_k(1 - \alpha_k) v_k - \alpha_k g_k + \mu \alpha_k y_k), \label{eqn:stepwise-wapg-eqn4}
            \\
            x_{k + 1} &= T_L y_k. \label{eqn:stepwise-wapg-eqn5}
        \end{align}
    \end{definition}
    % \begin{observation}\label{obs:stepwise-wapg}
    %     Let's observe Definition \ref{def:stepwise-wapg} closely. 
    %     We make two crucial observations here.     
    % \end{observation}

@@

        The proof starts by showing that \eqref{eqn:Q1}, \eqref{eqn:Q2} follows from \ref{eqn:stepwise-wapg-eqn2}, then the desire results follows naturally using algebra.
        We make the following observations from Definition \ref{def:stepwise-wapg}: 
            y_k - v_k &= 
            \frac{\hat \gamma_{k + 1}}{\alpha_k \gamma_k}(x_k - y_k), 
            \tag{Q1}\label{eqn:Q1}
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
            \tag{Q2}\label{eqn:Q2}
        It says that $x_k, y_k, v_k$ lie on the same line because \eqref{eqn:Q1} indicates $y_k - v_k$ parallels to $-(y_k - x_k)$ and both vector shares the same head which anchors at $y_k$. 
        To see \eqref{eqn:Q1}, we show that it's equivalent to \eqref{eqn:stepwise-wapg-eqn2} by considering: 
            y_k - v_k &= 
            -(\alpha_k \gamma_k \hat \gamma^{-1}_{k + 1} + 1)y_k
            - \alpha_k \gamma_k \hat \gamma^{-1}_{k + 1}v_k - x_k
            \frac{
                \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}v_k + x_k
            }{1 + \alpha_k \gamma_k \hat \gamma_{k + 1}^{-1}}
            \frac{\alpha_k \gamma_k v_k + \hat \gamma_{k + 1} x_k}{\gamma_k + \alpha_k \mu}.
        On the first equality \eqref{eqn:Q1}, we multiplied both side of the equation by $\hat\gamma_{k + 1}/\alpha_k \gamma_k$ and then grouped $y_k$. 
        The last equality comes by multiplying both the numerator and denominator by $\hat \gamma_{k + 1}$, then simplifies the denominator by rearranging \eqref{eqn:stepwise-wapg-eqn1} into $\hat \gamma_{k + 1} + \alpha_k \gamma_k = \gamma_k + \alpha_k \mu$. 
        To see \eqref{eqn:Q2}, consider \eqref{eqn:stepwise-wapg-eqn2}: 
            y_k &= (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \hat\gamma_{k + 1} x_k)
            y_k - x_k &= 
            (\alpha_k \gamma_k v_k - (\gamma_k + \alpha_k \mu)x_k + \hat \gamma_{k + 1} x_k)
            (\gamma_k + \alpha_k \mu)(y_k - x_k)
            \alpha_k\gamma _kv_k + 
            (\hat \gamma_k - \gamma_k - \alpha_k \mu) x_k
            &= \alpha_k \gamma_k v_k - \alpha_k \gamma_k x_k 
            &= \alpha_k \gamma_k(v_k - x_k)
            y_k - x_k &= 
            \frac{\alpha_k \gamma_k}{\gamma_k + \alpha_k \mu}(v_k - x_k). 
        On the second equality that follows, we substituted $\hat\gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \alpha_k \mu$ given by \eqref{eqn:stepwise-wapg-eqn1}. 
        Finally, we start by considering the LHS of \eqref{eqn:Q3} and it follows that: 

@@

        % PROOF SUMMARIZATIONS STARTS ==================================================================================
        Proceeding stepwise, we show the following intermediate results: 
        \begin{align}\tag{Step 1}\label{eqn:stepwise-lya-proof-step-1}
        \begin{align}\tag{Step 2}\label{eqn:stepwise-lya-proof-step-2}
                    + \langle v_k - x^*, \mu\alpha_k(y_k - v_k)\rangle,     
        \begin{align}\tag{Step 3}\label{eqn:stepwise-lya-proof-step-3}
                F(x_{k + 1}) + R_{k + 1} + 
        % \par 
        % We started with $F(x_{k + 1}) + R_{k + 1} \iff \eqref{eqn:stepwise-lya-proof-eqn1}$, 
        % then $\hat \gamma/2\Vert v_{k + 1} - x^*\Vert^2 = \eqref{expr:stepwise-lya-expr2.2}$ via \eqref{eqn:stepwise-lya-proof-eqn2}. 
        % We add (\ref{eqn:stepwise-lya-proof-eqn1}), (\ref{expr:stepwise-lya-expr2.2}) to get \eqref{eqn:stepwise-lya-proof-eqn3.1}. 
        % Finally, the entire LHS of \eqref{ineq:stepwise-lya-2} equals to the RHS of \eqref{eqn:stepwise-lya-proof-eqn3.2}. 
        % Going from \eqref{eqn:stepwise-lya-proof-eqn3.2} to \eqref{ineq:stepwise-lya-proof-ineq4}, we used Theorem \ref{thm:prox-grad-ineq} which yields the desired inequality.

@@

        \begin{align}\label{eqn:stepwise-lya-proof-eqn2}\begin{split}

@@

        \end{split}\end{align}
